---
title: "PIRLS Hypotheses--Reproducible Code"
author: 
date: "4/15/2022"
output:
  github_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


# Concatenate the data

Start by importing the data which has been downloaded from the PIRLS website. Then re-code the dependent variables.



```{r, eval = FALSE}
setwd("C:\\path\\ goes \\here")
pirls <-  intsvy::pirls.select.merge(folder = getwd(), 
                                     countries = c("USA"),
                                      student = c("IDSTUD"),
                                     school = c("ACBG03B"),
                                       teacher = c("ATBG05BB", "ATBG05BF", "ATBG05BC", "ATBR11F","ATBR13A", 
                                                   "ATBR20B", "ATBR11D", "ATBG05BF", "ATBG02", "ATBR12B", "ATBR11G", 
                                                   "ATBG05AA", "ATBR06","ATBR07","ATBR18","ATBR19A","ATBR19B","ATBR19C",
                                                   "ATBR11E","ATBR12E", "ATBR12C", "ATBG05BD", "ATBG10B", "ATBG06", "ATBR11E",
                                                   "ATBR12F", "ATBG05AD","ATBG05AB","ATBG05AA", "ATBR11C", "ATBR06", "ATBR07",
                                                   "ATBR14CA", "ATBR14CB", "ATBR14CC", "ATBR14CD","ATBR14CE", "ATBR14CF", "ATBG01","ATBR10D", "ATBG07A", "ATBR10B", "ATBG05AA"))


pirls$ATBR11F <- car::recode(pirls$ATBR11F, '1 = 4; 2 = 3; 3 = 2; 4 = 1') 

pirls$ATBR11G <- car::recode(pirls$ATBR11G, '1 = 4; 2 = 3; 3 = 2; 4 = 1') 

pirls$ATBR12B <- car::recode(pirls$ATBR12B, '1 = 4; 2 = 3; 3 = 2; 4 = 1') 

pirls$ATBR11E <- car::recode(pirls$ATBR11E, '1 = 4; 2 = 3; 3 = 2; 4 = 1') 

pirls$ATBR12C <- car::recode(pirls$ATBR12C, '1 = 4; 2 = 3; 3 = 2; 4 = 1')

pirls$ATBR13A <- car::recode(pirls$ATBR13A, '1 = 4; 2 = 3; 3 = 2; 4 = 1') 

pirls$ATBR11D <- car::recode(pirls$ATBR11D, '1 = 4; 2 = 3; 3 = 2; 4 = 1') 

pirls$ATBR12B <- car::recode(pirls$ATBR12B, '1 = 4; 2 = 3; 3 = 2; 4 = 1') 

pirls$ATBR11G <- car::recode(pirls$ATBR11G, '1 = 4; 2 = 3; 3 = 2; 4 = 1') 

pirls$ATBR18 <- car::recode(pirls$ATBR18, '1 = 4; 2 = 3; 3 = 2; 4 = 1') 

pirls$ATBR19A <- car::recode(pirls$ATBR19A, '1 = 3; 2 = 3; 3 = 2') # only goes up to 3

pirls$ATBR19B <- car::recode(pirls$ATBR19B, '1 = 3; 2 = 3; 3 = 2') # only goes up to 3

pirls$ATBR19C <- car::recode(pirls$ATBR19C, '1 = 3; 2 = 3; 3 = 2') # only goes up to 3

pirls$ATBR11E <- car::recode(pirls$ATBR11E, '1 = 4; 2 = 3; 3 = 2; 4 = 1') 

pirls$ATBR12E <- car::recode(pirls$ATBR12E, '1 = 4; 2 = 3; 3 = 2; 4 = 1') 

pirls$ATBR12C <- car::recode(pirls$ATBR12C, '1 = 4; 2 = 3; 3 = 2; 4 = 1') 

pirls$ATBR11E <- car::recode(pirls$ATBR11E, '1 = 4; 2 = 3; 3 = 2; 4 = 1') 

pirls$ATBR12F <- car::recode(pirls$ATBR12F, '1 = 4; 2 = 3; 3 = 2; 4 = 1') 

pirls$ATBR11D <- car::recode(pirls$ATBR11D, '1 = 4; 2 = 3; 3 = 2; 4 = 1') 

pirls$ATBR11G <- car::recode(pirls$ATBR11G, '1 = 4; 2 = 3; 3 = 2; 4 = 1') 

pirls$ATBR11C <- car::recode(pirls$ATBR11C, '1 = 4; 2 = 3; 3 = 2; 4 = 1') 

pirls$ATBR14CA <- car::recode(pirls$ATBR14CA, '1 = 4; 2 = 3; 3 = 2; 4 = 1') 

pirls$ATBR14CB <- car::recode(pirls$ATBR14CB, '1 = 4; 2 = 3; 3 = 2; 4 = 1') 

pirls$ATBR14CC <- car::recode(pirls$ATBR14CC, '1 = 4; 2 = 3; 3 = 2; 4 = 1') 

pirls$ATBR14CD <- car::recode(pirls$ATBR14CD, '1 = 4; 2 = 3; 3 = 2; 4 = 1') 

pirls$ATBR14CE <- car::recode(pirls$ATBR14CE, '1 = 4; 2 = 3; 3 = 2; 4 = 1') 

pirls$ATBR14CF <- car::recode(pirls$ATBR14CF, '1 = 4; 2 = 3; 3 = 2; 4 = 1') 

pirls$ATBR10D <- car::recode(pirls$ATBR10D, '1 = 4; 2 = 3; 3 = 2; 4 = 1') 

pirls$ATBG07A <- car::recode(pirls$ATBG07A, '1 = 4; 2 = 3; 3 = 2; 4 = 1') 

pirls$ATBR10B <- car::recode(pirls$ATBR10B, '1 = 4; 2 = 3; 3 = 2; 4 = 1') 



```

# Select variables

and sample the data using dplyr. Save the sampled data as an object. 

```{r, eval = F}

library(dplyr)

pirls1 <- pirls %>%
  dplyr::select(ATBG05AD, ATBG07A, ATBG01,ATBR11D, ATBR12E,ATBR11G, ATBR11F,ATBR11C, ATBR11E, ATBR14CA, ATBR14CC, ATBR14CC, ATBR14CD, ATBR14CE, ATBR14CF, ATBR06, ATBR07, ATBR10D, ATBR10B,ATBG05AA)

pirls2 <- na.omit(pirls1) # omit rows with missing information

Samp <- pirls2 %>%
  group_by(ATBG05AA) %>%
  dplyr::filter(., ATBG01 < 6) #%>%
 # sample_frac(.9) # sample the dataframe...actually not needed



```
## Work with the data

to run tests. Shapiro test in R to run for normality,

```{r, eval = F}

setwd("C:\\Users\\mario\\OneDrive\\Desktop\\PIRLS\\P16_SPSSData_pt2")
pirls <-  intsvy::pirls.select.merge(folder = getwd(), 
                                     countries = c("USA"),
                                      student = c("IDSTUD"),
                                     school = c("ACBG03B"),
                                       teacher = c("ATBG05BB", "ATBG05BF", "ATBG05BC", "ATBR11F","ATBR13A", 
                                                   "ATBR20B", "ATBR11D", "ATBG05BF", "ATBG02", "ATBR12B", "ATBR11G", 
                                                   "ATBG05AA", "ATBR06","ATBR07","ATBR18","ATBR19A","ATBR19B","ATBR19C",
                                                   "ATBR11E","ATBR12E", "ATBR12C", "ATBG05BD", "ATBG10B", "ATBG06", "ATBR11E",
                                                   "ATBR12F", "ATBG05AD","ATBG05AB","ATBG05AA", "ATBR11C", "ATBR06", "ATBR07",
                                                   "ATBR14CA", "ATBR14CB", "ATBR14CC", "ATBR14CD","ATBR14CE", "ATBR14CF", "ATBG01","ATBR10D", "ATBG07A", "ATBR10B", "ATBG05AA"))


pirls$ATBR11F <- car::recode(pirls$ATBR11F, '1 = 4; 2 = 3; 3 = 2; 4 = 1') 

pirls$ATBR11G <- car::recode(pirls$ATBR11G, '1 = 4; 2 = 3; 3 = 2; 4 = 1') 

pirls$ATBR12B <- car::recode(pirls$ATBR12B, '1 = 4; 2 = 3; 3 = 2; 4 = 1') 

pirls$ATBR11E <- car::recode(pirls$ATBR11E, '1 = 4; 2 = 3; 3 = 2; 4 = 1') 

pirls$ATBR12C <- car::recode(pirls$ATBR12C, '1 = 4; 2 = 3; 3 = 2; 4 = 1')

pirls$ATBR13A <- car::recode(pirls$ATBR13A, '1 = 4; 2 = 3; 3 = 2; 4 = 1') 

pirls$ATBR11D <- car::recode(pirls$ATBR11D, '1 = 4; 2 = 3; 3 = 2; 4 = 1') 

pirls$ATBR12B <- car::recode(pirls$ATBR12B, '1 = 4; 2 = 3; 3 = 2; 4 = 1') 

pirls$ATBR11G <- car::recode(pirls$ATBR11G, '1 = 4; 2 = 3; 3 = 2; 4 = 1') 

pirls$ATBR18 <- car::recode(pirls$ATBR18, '1 = 4; 2 = 3; 3 = 2; 4 = 1') 

pirls$ATBR19A <- car::recode(pirls$ATBR19A, '1 = 3; 2 = 3; 3 = 2') # only goes up to 3

pirls$ATBR19B <- car::recode(pirls$ATBR19B, '1 = 3; 2 = 3; 3 = 2') # only goes up to 3

pirls$ATBR19C <- car::recode(pirls$ATBR19C, '1 = 3; 2 = 3; 3 = 2') # only goes up to 3

pirls$ATBR11E <- car::recode(pirls$ATBR11E, '1 = 4; 2 = 3; 3 = 2; 4 = 1') 

pirls$ATBR12E <- car::recode(pirls$ATBR12E, '1 = 4; 2 = 3; 3 = 2; 4 = 1') 

pirls$ATBR12C <- car::recode(pirls$ATBR12C, '1 = 4; 2 = 3; 3 = 2; 4 = 1') 

pirls$ATBR11E <- car::recode(pirls$ATBR11E, '1 = 4; 2 = 3; 3 = 2; 4 = 1') 

pirls$ATBR12F <- car::recode(pirls$ATBR12F, '1 = 4; 2 = 3; 3 = 2; 4 = 1') 

pirls$ATBR11D <- car::recode(pirls$ATBR11D, '1 = 4; 2 = 3; 3 = 2; 4 = 1') 

pirls$ATBR11G <- car::recode(pirls$ATBR11G, '1 = 4; 2 = 3; 3 = 2; 4 = 1') 

pirls$ATBR11C <- car::recode(pirls$ATBR11C, '1 = 4; 2 = 3; 3 = 2; 4 = 1') 

pirls$ATBR14CA <- car::recode(pirls$ATBR14CA, '1 = 4; 2 = 3; 3 = 2; 4 = 1') 

pirls$ATBR14CB <- car::recode(pirls$ATBR14CB, '1 = 4; 2 = 3; 3 = 2; 4 = 1') 

pirls$ATBR14CC <- car::recode(pirls$ATBR14CC, '1 = 4; 2 = 3; 3 = 2; 4 = 1') 

pirls$ATBR14CD <- car::recode(pirls$ATBR14CD, '1 = 4; 2 = 3; 3 = 2; 4 = 1') 

pirls$ATBR14CE <- car::recode(pirls$ATBR14CE, '1 = 4; 2 = 3; 3 = 2; 4 = 1') 

pirls$ATBR14CF <- car::recode(pirls$ATBR14CF, '1 = 4; 2 = 3; 3 = 2; 4 = 1') 

pirls$ATBR10D <- car::recode(pirls$ATBR10D, '1 = 4; 2 = 3; 3 = 2; 4 = 1') 

pirls$ATBG07A <- car::recode(pirls$ATBG07A, '1 = 4; 2 = 3; 3 = 2; 4 = 1') 

pirls$ATBR10B <- car::recode(pirls$ATBR10B, '1 = 4; 2 = 3; 3 = 2; 4 = 1') 



```

# Select variables, and sample the data using dplyr. Save the sampled data as an object. 

```{r, eval = F}

library(tidyverse)

# Strategy 1: sample the whole group

library(dplyr)

pirls1 <- pirls %>%
  dplyr::select(ATBG05AD, ATBG07A, ATBG01,ATBR11D, ATBR12E,ATBR11G, ATBR11F,ATBR11C, ATBR11E, ATBR14CA, ATBR14CC, ATBR14CC, ATBR14CD, ATBR14CE, ATBR14CF, ATBR06, ATBR07, ATBR10D, ATBR10B,ATBG05AA)

pirls2 <- pirls1 %>%
  tidyr::drop_na(ATBG01) # omit rows with missing information



Samp <- pirls2 %>%
  group_by(ATBG05AA) %>%
  dplyr::filter(., ATBG01 < 6) 

# n = 922 

saveRDS(Samp, "c:/users/mario/OneDrive/Desktop/PIRLS/SAMPLED DATA/beginningteachers_correct_variable_5yrs-922.rds")


```


## Work with the data to run tests. Shaprio-wilks to run for normality, Wilcoxon tests to run for non-parametric version of t-tests, and the t-test

```{r}

library(tidyverse)
# ATBR12E ATBR11D, ATBR12E,ATBR11G, ATBR11F,ATBR11C,ATBR11C, ATBR11E, ATBR14CA, ATBR14CC,  ATBR14CD, ATBR14CE, ATBR14CF, ATBR06, ATBR07, ATBR10D



mydata <- readRDS("c:/users/mario/OneDrive/Desktop/PIRLS/SAMPLED DATA/beginningteachers_correct_variable_5yrs-922.rds")

mydata$ATBG05AA <- as.character(mydata$ATBG05AA)

```


```{r}

mydata[1:18] %>%
map(., function(x) shapiro.test(x)) 

```

```{r}

lapply(mydata[-19], function(x) WRS2::yuenbt(x ~ mydata$ATBG05AA))

```

```{r}
psych::describeBy(mydata, mydata$ATBG05AA) # get descriptive statistics

```

```{r}

lapply(mydata[-19], function(x) WRS2::akp.effect(x ~ mydata$ATBG05AA))


```


